### 2017-06-21 paper/jfp-submission/evaluation.tex
update some numbers, still need investigation of failed witness

\subsection{Methodology}
\label{sec:methodology}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs
from a cohort of 46 students..
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
and polymorphic and algebraic data types. % and expression evaluators.

@@

\begin{figure}[t]
% \centerline{
% \begin{minipage}{1.2\textwidth}
\centering
\includegraphics[width=0.7\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
% \includegraphics[width=0.6\linewidth]{distrib.png}
% \end{minipage}
% }
% \vspace{3ex}
\caption{Results of our coverage testing. Our random search successfully
  finds witnesses for 76--82\% of the programs in under one second,
  improving to 84--85\% in under 10 seconds. }
\label{fig:results-witness}
\end{figure}
\begin{figure}[t]
\includegraphics[width=0.7\linewidth]{distrib.png}
\caption{Distribution of test outcomes. In both datasets we detect
  actual type errors at least 75\% of the time, unbound variables or
  constructors 4--5\% of the time, and diverging loops 3\% of the
  time. For the remaining 13--16\% of the programs we are unable to
  provide any useful feedback. }
\label{fig:results-distrib}
\end{figure}

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figures~\ref{fig:results-witness}~and~\ref{fig:results-distrib}.
%
In both datasets our tool was able to find a witness for over 75\% of the
programs in under one second, \ie\ fast enough to be integrated as a
compile-time check. If we extend our tolerance to a 10 second timeout,
we reach 84--85\% coverage, and if we allow a 60 second search,
we hit a maximum of 84--87\% coverage.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 4--5\% triggered an unbound variable error (even
though \ocaml\ reported a type error) and 3\% triggered an infinite
recursion error.
%
For the remaining 13--16\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after one minute.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg\ dynamic-symbolic execution,
could likely uncover more errors, our experiments suggest that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

@@

% \item \ES{others?}
%
\begin{figure}[t]
% \centerline{
% \begin{minipage}{1.2\textwidth}
\includegraphics[width=0.7\linewidth]{trace_size_step.png}
\includegraphics[width=0.7\linewidth]{trace_size_jump.png}
% \end{minipage}
% }
% \vspace{3ex}
\caption{Complexity of the generated traces. Over 80\% of the combined traces
  have a jump complexity of at most 10, with an average complexity of 7
  and a median of 5.}
\label{fig:results-complexity}
\end{figure}
%
\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 30 for the
\ucsdbench\ dataset (45 for the \uwbench\ dataset) with a maximum of
2,745 (\resp 982) and a median of 14.5 (\resp 14.5).
%
The average number of jumps per trace is 7 (\resp 10) with a
maximium of 353 (\resp 221) and a median of 4 (\resp 4).
%
In both datasets 60\% of traces have at most 5 jumps, and 80\% or more
have at most 10 jumps.

>>>>

\subsection{Methodology}
\label{sec:methodology}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs
from a cohort of 46 students.
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
and polymorphic and algebraic data types. % and expression evaluators.

@@

\begin{figure}[t]
% \centerline{
% \begin{minipage}{1.2\textwidth}
\centering
\includegraphics[width=0.7\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
% \includegraphics[width=0.6\linewidth]{distrib.png}
% \end{minipage}
% }
% \vspace{3ex}
\caption{Results of our coverage testing. Our random search successfully
  finds witnesses for 76--83\% of the programs in under one second,
  improving to 84--85\% in under 10 seconds. }
\label{fig:results-witness}
\end{figure}
\begin{figure}[t]
\includegraphics[width=0.7\linewidth]{distrib.png}
\caption{Distribution of test outcomes. In both datasets we detect
  actual type errors at least 77\% of the time, unbound variables or
  constructors 4\% of the time, and diverging loops 2--3\% of the
  time. For the remaining 15--16\% of the programs we are unable to
  provide any useful feedback. }
\label{fig:results-distrib}
\end{figure}

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figures~\ref{fig:results-witness}~and~\ref{fig:results-distrib}.
%
In both datasets our tool was able to find a witness for over 75\% of the
programs in under one second, \ie\ fast enough to be integrated as a
compile-time check. If we extend our tolerance to a 10 second timeout,
we reach 84\% coverage, and if we allow a 60 second search,
we hit a maximum of 84--85\% coverage.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 4\% triggered an unbound variable error (even
though \ocaml\ reported a type error) and 3\% triggered an infinite
recursion error.
%
For the remaining 15--16\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after one minute.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg\ dynamic-symbolic execution,
could likely uncover more errors, our experiments suggest that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

@@

% \item \ES{others?}
%
\begin{figure}[t]
% \centerline{
% \begin{minipage}{1.2\textwidth}
\includegraphics[width=0.7\linewidth]{trace_size_step.png}
\includegraphics[width=0.7\linewidth]{trace_size_jump.png}
% \end{minipage}
% }
% \vspace{3ex}
\caption{Complexity of the generated traces. Over 80\% of the combined traces
  have a jump complexity of at most 10, with an average complexity of 7
  and a median of 5.}
\label{fig:results-complexity}
\end{figure}
%
\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 17 for the
\ucsdbench\ dataset (42 for the \uwbench\ dataset) with a maximum of
2,745 (\resp 982) and a median of 15 (\resp 15).
%
The average number of jumps per trace is 7 (\resp 9) with a
maximium of 353 (\resp 221) and a median of 4 (\resp 4).
%
In both datasets about 60\% of traces have at most 5 jumps, and 80\% or more
have at most 10 jumps.

### 2017-06-21 paper/jfp-submission/locating.tex
update some numbers, still need investigation of failed witness

Not all ill-typed programs will have an associated fix; furthermore,
at some point a ``fix'' becomes a ``rewrite''.
%
We do not wish to consider the ``rewrites'', so we discard outliers
where the fraction of expressions that have changed is more than one
standard deviation above the mean, establishing a diff threshold of
45\%.
%
This accounts for roughly 14\% of programs pairs we discovered, leaving
us with 2,425 program pairs.

@@

We evaluate each tool based on whether \emph{any} of its predictions
identifies a changed expression.
%
There were a number of programs where \mycroft or \sherrloc
encountered an unsupported language feature or timed out after one
minute, or where \toolname failed to produce a witness.
%
We discard all such programs in our evaluation to level the playing
field, around 15\% for each tool, leaving us with a benchmark set of
1,569 programs.

\paragraph{Threats to Validity}
Our benchmarks were drawn from students in an undergraduate course
at \ucsdbench\ and may not be representative of other student bodies.
%
We mitigate this threat with a large empirical evaluation of 1,569
programs, drawn from a cohort of 46 students.
%
A similar threat is that students are not industrial programmers, thus
our results may not translate to large-scale software engineering.
%
% In particular, \toolname's reliance on running the program suggests that
% scaling it to larger programs would be more difficult than static
% methods.
%
However, in our experience programmers are able to construct a mental
model of type systems after sufficient exposure, at which point
traditional error reports may suffice.
%
We are thus particularly interested in aiding novice programmers as
they learn to work with the type system.

@@

\begin{figure}[t]
\includegraphics[width=0.7\linewidth]{blame.png}
\caption{Accuracy of type error localization. \toolname's witness-based
  predictions outperform \ocaml by 22 points, and are competitive
  with the state-of-the-art tools \mycroft and \sherrloc.}
\label{fig:results-blame}
\end{figure}

\paragraph{Results}
Figure~\ref{fig:results-blame} summarizes our results, which show that
\toolname's witnesses are competitive with \mycroft and \sherrloc in
automatically locating the source of a type error.
%
\toolname, \mycroft, and \sherrloc all outperform the \ocaml compiler,
which is not surprising given that they can produce multiple possible
error locations, while the \ocaml compiler is limited to one predicted
error location.
%
Interestingly, while all tools have a median of 2 predicted error
locations per program, \mycroft and \sherrloc have a long tail with a
maximum of 22 (\resp 12) locations, while \toolname's maximum is 5
locations.
%
We also note that while \mycroft and \sherrloc were designed
specifically to \emph{localize} type errors, \toolname's foremost
purpose is to \emph{explain} them, we consider its ability to localize
type errors an added benefit.

>>>>

Not all ill-typed programs will have an associated fix; furthermore,
at some point a ``fix'' becomes a ``rewrite''.
%
We do not wish to consider the ``rewrites'', so we discard outliers
where the fraction of expressions that have changed is more than one
standard deviation above the mean, establishing a diff threshold of
40\%.
%
This accounts for roughly 14\% of programs pairs we discovered, leaving
us with 2,710 program pairs.

@@

We evaluate each tool based on whether \emph{any} of its predictions
identifies a changed expression.
%
There were a number of programs where \mycroft or \sherrloc
encountered an unsupported language feature or timed out after one
minute, or where \toolname failed to produce a witness.
%
We discard all such programs in our evaluation to level the playing
field, around 15\% for each tool, leaving us with a benchmark set of
1,759 programs.

\paragraph{Threats to Validity}
Our benchmarks were drawn from students in an undergraduate course
at \ucsdbench\ and may not be representative of other student bodies.
%
We mitigate this threat with a large empirical evaluation of 1,759
programs, drawn from a cohort of 46 students.
%
A similar threat is that students are not industrial programmers, thus
our results may not translate to large-scale software engineering.
%
% In particular, \toolname's reliance on running the program suggests that
% scaling it to larger programs would be more difficult than static
% methods.
%
However, in our experience programmers are able to construct a mental
model of type systems after sufficient exposure, at which point
traditional error reports may suffice.
%
We are thus particularly interested in aiding novice programmers as
they learn to work with the type system.

@@

\begin{figure}[t]
\includegraphics[width=0.7\linewidth]{blame.png}
\caption{Accuracy of type error localization. \toolname's witness-based
  predictions outperform \ocaml by 21 points, and are competitive
  with the state-of-the-art tools \mycroft and \sherrloc.}
\label{fig:results-blame}
\end{figure}

\paragraph{Results}
Figure~\ref{fig:results-blame} summarizes our results, which show that
\toolname's witnesses are competitive with \mycroft and \sherrloc in
automatically locating the source of a type error.
%
\toolname, \mycroft, and \sherrloc all outperform the \ocaml compiler,
which is not surprising given that they can produce multiple possible
error locations, while the \ocaml compiler is limited to one predicted
error location.
%
Interestingly, while all tools have a median of 2 predicted error
locations per program, \mycroft and \sherrloc have a long tail with a
maximum of 22 (\resp 11) locations, while \toolname's maximum is 5
locations.
%
We also note that while \mycroft and \sherrloc were designed
specifically to \emph{localize} type errors, \toolname's foremost
purpose is to \emph{explain} them, we consider its ability to localize
type errors an added benefit.
